---
title: "607 Final Project"
author: "Dylan Gold"
format: html
editor: visual
---

# Approach
For this project I will be collecting data on video game sales as well as video game reviews and seeing how factors like review score and genre affect videogame sales.

# Collecting data
To start this project we need to first collect the data. \
I have chosen two data sources one is a csv from https://www.kaggle.com/datasets/volodymyrpivoshenko/video-game-sales-dataset \
The other is using an api of a gaming review website rawg.io. This is a link to their api documentation https://api.rawg.io/docs/#tag/developers \

First I will collect the data from the csv. This will serve as our base data that we will use to query the api.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
```


```{r, message=FALSE, warning=FALSE}
url <- "https://raw.githubusercontent.com/DylanGoldJ/607-Final-Project/refs/heads/main/video_games_sales.csv"

df <- read_csv(
  file = url,
  col_names = TRUE
)


head(df, 5)
```
Now I also need to collect data from the api. \
I will start with a simple example. \
First I will load in my key
```{r}
api_key <- Sys.getenv("RAWGIO_API_KEY") # Load api key from envrionment.
```

Because the api has way more total games than our data set we will query the data set for each game individually.\
First I will try to get Wii Sports game info from the api. \

```{r}
api_url <- "https://api.rawg.io/api/games"

response <- GET(
  api_url,
  query = list(
    key = api_key, # Api key
    search = "Wii Sports" # Wii sports as our testing value.
  )
)

status_code(response)
```
We got a sucessful response. Now I will extract the data from the response
```{r}
data <- fromJSON(content(response, "text", encoding = "UTF-8"))
game_review <- data$results
head(game_review)
```

We can see that we were given a list of games that match the search query. The top match Wii Sports is exactly what we searched for. \
We can also see that the second game from this search is also the 4th ranked game on our games sold data frame. \
There will be times when we get multiple games from a search query that are in our games sold data frame. \
To minimize the number of requests (I have only 20,000 this month) I will try to fill in this data as well from each search. \

This was a challenge I had, making sure to not waste my limited requests. Initially my plan was to search each game one at a time. While I may have had enough requests to do this it would be a lot slower as well as waste requests that I may need for the future. \
Because the search was not an exact match, we can get values that we did not search directly which helps maintain our query limit.

First I will make a function that takes the values from the request for the game and binds them to the original data frame with a left join.
```{r}
game_data <- df # Copy of df to work on

# Make new columns in game_data

# This function will take the sold df and the reviews df and fill in columns for the game_data df based on the df_reviews.
combine_data <- function(df, df_reviews){
  # Select columns from reviews that we want to add.
  head(df_reviews)
  df_reviews <- df_reviews %>% select( # Select different rating to use, `rating` seems to be the main one.
    Name = name,
    rating, 
    rating_top,
    metacritic,
    score,
    ratings_count
  )
  
  combined <- left_join(df, df_reviews, by = "Name")
  return(combined)
}
combined_wii <- combine_data(game_data, game_review)
head(combined_wii)

```
We can see that there are several columns that got filled in below is some of the values
```{r}
head(combined_wii %>% filter(!is.na(rating))) # Check rows with a rating
```

Now I will make a function that can get the api request data given the search term. \
It will return the data frame with the data or null if the request is bad

```{r}
search_api <- function(query){

api_url <- "https://api.rawg.io/api/games"

response <- GET(
  api_url,
  query = list(
    key = api_key, # Api key
    search = query # Wii sports as our testing value.
  )
)

if (status_code(response) == 200){
  data <- fromJSON(content(response, "text", encoding = "UTF-8"))
  game_review <- data$results
  return(game_review)
  }
return(NULL) # If we have a bad response return null
}
```

I will test this function on a popular series Call of Duty. I will search for a specific game "Call of Duty: Black Ops" \
```{r}
cod_response_df <- search_api("Call of Duty: Black Ops")
head(cod_response_df)
```
I wanted to look at how this game shows up in our combined data.
```{r}
combined_data <- combined_wii

cod_sales <- combined_data %>% 
  filter(str_detect(Name, "Call of Duty"))
head(cod_sales)
```

We can combine these like we did before with the Wii sports games. \
This time the columns are already created so we have to also coalesce the data frames.\
I will update the function I made to do this. \
If we were doing it one by one we could combine it all in one step but I wanted a way that made the data frame step by step.

```{r}
# Update the function to handle the columns already being present.
combine_data <- function(df, df_reviews){
  # Select columns from reviews that we want to add.
  head(df_reviews)
  df_reviews <- df_reviews %>% select( # Select different rating to use, `rating` seems to be the main one.
    Name = name,
    rating, 
    rating_top,
    metacritic,
    score,
    ratings_count
  )
  
  combined <- left_join(df, df_reviews, by = "Name") %>%
  mutate(
    rating = coalesce(rating.x, rating.y), # Because the dataframe already has these columns we can coalesce. This will take the non na value.
    rating_top = coalesce(rating_top.x, rating_top.y),
    metacritic = coalesce(metacritic.x, metacritic.y),
    score = coalesce(score.x, score.y),
    ratings_count = coalesce(ratings_count.x, ratings_count.y)
    ) %>%
  select(-ends_with(".x"), -ends_with(".y")) # Remove the .x and .y columns. These are the duplicates from the original left_join 
  return(combined)
}

combined_data <- combine_data(combined_data, cod_response_df)
(combined_data %>% filter(!is.na(rating) &  (str_detect(Name, "Call of Duty"))))  # Show the cod games that were filled
```
While looking at the data I notice that not all of the Call of Duty games from the reviews have matches to our sales data.\
It is entirely possible that these data sets do not have complete overlap. I will leave these values as NA. If we have enough data I will leave it like this. \

Now I want to iterate through the data frame based on null ratings. \
I used the latest free version of chat gpt to help generate this code. \
This code keeps track of the already tried rows
```{r, cache = TRUE}
attempted_ids <- character() # Keep track of attempted names
```

```{r, cache = TRUE}

remaining_ids <- combined_data %>% # Get the names that still need to be tried
  filter(is.na(rating)) %>% # Rows that have no rating
  distinct(Name) # No need to run the same game twice(its separated by Platform)

while(nrow(remaining_ids) > 0) { # while we have remaining ids

  id <- remaining_ids$Name[1] # Get the name as id

  # mark as attempted BEFORE api call
  attempted_ids <- c(attempted_ids, id) # Update attempted_ids with the extra name
  
  api_result <- search_api(id) # Search for the name
  print(id)
  # update data frame
  if (!is.null(api_result)) {
    if(length(api_result) != 0) # Check empty/na api_result
      combined_data <- combine_data(combined_data, api_result)
}

  # only unresolved AND unattempted ids 
  remaining_ids <- combined_data %>% 
    filter(is.na(rating)) %>%
    distinct(Name) %>% 
    filter(!Name %in% attempted_ids) #Names that are not attempted
}
```
Now that I have the data, because I don't want to have to run the api request again to knit the project, and I still want the data to be accessible and reproducible I will save the data to a csv. 

```{r}
write.csv(df, "combined_data.csv", row.names = TRUE)
```


The most difficult part was getting the data, now that we have the data we can create graphs and such to try to find a pattern between the games sold and 
```{r}
combined_by_name <- combined_data %>%
  group_by(Name) %>%
  summarize()
```
